---
title: "DATA607-Week7"
output: html_notebook
---

In addition to being a fun and educational exercise in and of itself, I wanted to use the exercise to answer an interest question (well, at least one interesting to me!). So I explored a few of the various APIs that NYT makes available, and many presented potentially interesting ideas. In the end, maybe because I had been struggling to find a good novel to read, I ended up choosing the books API. But I didn't just want to pull down a list of best sellers. I wanted to dig into the data just a bit and see if I could uncover something interesting. I'd like to know whether Fiction books stay on the best sellers list longer than Non-Fiction books or vice versa. At this point, I've only confirmed feasibility of what I'd like to do, but I haven't actually captured any data, so I don't know what the answer is yet.  

According to the NYT API README page, the best sellers list URI format is as follows:  
`http://api.nytimes.com/svc/books/{version}/lists/[date/]{list-name}[.response_format]?`
`[optional-param1=value1]&[...]&api-key={your-API-key}`   

```{r message=FALSE, warning=FALSE}

library(jsonlite)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggthemes)

api_key = 'api-key=368fbac19ccc4179ae2db761a426699d'
base_url = 'https://api.nytimes.com/svc/books/v3/lists/'
resp_format = '.json'

fict = 'combined-print-and-e-book-fiction'
nonfict = 'combined-print-and-e-book-nonfiction'

fict_url = paste0(base_url, fict, resp_format, '?', api_key)
fict_res = fromJSON(fict_url)
fict_df = fict_res$results$books
fict_df$category = 'fiction'

nonfict_url = paste0(base_url, nonfict, resp_format, '?', api_key)
nonfict_res = fromJSON(nonfict_url)
nonfict_df = nonfict_res$results$books
nonfict_df$category = 'nonfiction'

nytbestsellers = bind_rows(fict_df, nonfict_df)

nytbestsellers$vendor_link = paste0('[Amazon](', nytbestsellers$amazon_product_url, ')')
nytbestsellers$cover = paste0("<img src=", nytbestsellers$book_image, " width='50px'/>")
```

Let's first have a look at the top five fiction and non-fiction books currently on the list.  

```{r}
knitr::kable(nytbestsellers[1:5, c('category', 'rank', 'title', 'description', 'author', 'vendor_link', 'cover')], row.names = FALSE, caption = 'New York Times Best Sellers - Top 5 in Fiction') 

knitr::kable(nytbestsellers[21:25, c('category', 'rank', 'title', 'description', 'author', 'vendor_link', 'cover')], row.names = FALSE, caption = 'New York Times Best Sellers - Top 5 in Non-Fiction')
```

Next, we'll take a look at how long books have been on their respective best sellers lists. Based on the histograms below, there doesn't seem to be a discernible difference between the two book categories.  

```{r}
nytbestsellers %>% ggplot(aes(x = weeks_on_list, col = category, fill = category)) +
  geom_histogram(alpha = 0.5, binwidth = 1, position = 'identity') +
  facet_grid(category ~ .) + 
  scale_fill_economist() + scale_color_economist() + 
  ggtitle('Number of Weeks on NYT Best Sellers List Histogram')
```

Perhaps the amount of turnover within the top 20 of each category might prove more interesting. Let's look at how many new books were added to each list that weren't there the previous list. Fortunately, the API results provide each book's ranking for the previous week. All we need to do is count how many were not in the top 20 in the previous week.  

And again, there does not seem to be a significant difference. There were 10 new non-fiction additions compared to 9 new fiction additions. 

```{r}
nytbestsellers %>% mutate(newAddition = ifelse(rank_last_week == 0, 'New Addition', 'Not New Addition')) %>%
  group_by(category, newAddition) %>%
  summarise(n = n()) %>% 
  mutate(freq = n / sum(n)) %>%
  ggplot(aes(x = category, y = freq, fill = newAddition)) +
  geom_bar(stat = 'identity') +
  scale_fill_economist() + scale_color_economist() + 
  ggtitle('Proportion of New Additions to NYT Best Sellers List')
```

I'm running out of things to compare here! One last comparison then - let's have a look at how much movement in rankings took place within each category from the previous week to the current week. We'll evaluate the change in rank from week-to-week. For those books that were not ranked in the previous week, we'll assume that they were ranked 21 (just outside of the best sellers list).  

Alas, when comparing the average absolute value of rank change, fiction books seem to have moved around just a bit more than non-fiction books; however, without a benchmark, it's hard to say how big of a difference this really represents.  

```{r}
nytbestsellers %>% mutate(rankChange = ifelse(rank_last_week == 0, 21, rank_last_week) - rank) %>%
  group_by(category) %>%
  summarise(avgAbsRankChange = mean(abs(rankChange)))
```

All in all, this was a fun exercise even if the takeaways from the data analysis were not groundbreaking. Perhaps looking at this data over time would prove more interesting...  


