---
title: "Hao-HW2"
author: "Bruce Hao"
date: "September 7, 2016"
output: pdf_document
---

Setup.
```{r message=FALSE, warning=FALSE}
library(VennDiagram)
```


**2.6**

```{r}
combinations = expand.grid(seq(1,6), seq(1,6))
combinations$Total = combinations$Var1 + combinations$Var2
nrow(combinations[combinations$Total==5, ])
```

a) 0
b) 4/36 = 11.11%
c) 1/36 = 2.78%


**2.8**

a) No, disjoint outcomes are mutually exclusive.
b)

```{r fig.align='center'}
draw.pairwise.venn(area1 = 14.6, area2 = 20.7, cross.area = 4.2, 
                   category = c('Americans in poverty %', 'Foreign language speakers at home %'), 
                   cat.pos = c(3, -1))
```

c) 10.4%
d) 16.5 + 4.2 + 10.4 = 31.1%
e) 100 - 31.1 = 68.9%
f) No, the probability of being below the poverty line is 14.6% for all Americans, but the probability of being below the poverty line given being a foreign language speaker at home is 4.2/20.7 = 20.29%


**2.20**

a) (114 + 108) / 408 = 54.41%
b) 78 / 114 = 68.42%
c) 19 / 54 = 35.19%; 11 / 36 = 30.56%
d) No, the color of the male's eyes affects the probability of having a partner with certain eye color.


**2.30**

a) 28/95 * 59/94 = 18.50%
b) 67/95 * 28/94 = 21.01%
c) 67/95 * 28/95 = 20.79%
d) When the sample size is a small fraction of the population, observations are nearly independent even when sampling without replacement. 


**2.38**

a) 0 * 0.54 + 25 * 0.34 + 35 * 0.12 = $12.7 average revenue per passenger;
  0.54*(0 - 12.7)^2 + 0.34*(25 - 12.7)^2 + 0.12*(35 - 12.7)^2 = $198.21 variance;
  198.21^0.5 = $14.08 standard deviation
b) $12.7 * 120 = $1,524 revenue per flight; 
  ($198.21 * 120)^0.5 = $154 standard deviation


**2.44**

a) The distribution is bell shaped with a minimum of zero and a fat right tail.
b) 62.2%
c) 62.2% * 41% = 25.5%; this assumes that gender and income are independent variables.
d) Based on the table below, 29.4% of femals earn less than $50K per year; thus, the assumption in part c is likely incorrect (but further analysis would be needed to answer this more concretely)
```{r echo=FALSE}
table = matrix(c(32.8, 29.4, 62.2, 26.2, 11.6, 37.8, 59.0, 41.0, 100.0), ncol=3, byrow=T)
colnames(table) = c('M', 'F', 'Total')
rownames(table) = c('<$50K', '>=$50K', 'Total')
table
```

