---
title: "Hao-2_7"
author: "Bruce Hao"
date: "September 8, 2016"
output: html_document
---

Setup.
```{r message=FALSE, warning=FALSE}
library(VennDiagram)
```


**2.7**

Proportional contingency table: 

```{r echo=FALSE}
table = matrix(c(11,24,35,12,53,65,23,77,100), ncol=3, byrow=T)
colnames(table) = c('Swing', 'Not Swing', 'Total')
rownames(table) = c('Independent', 'Not Independent', 'Total')
table
```

a) Are being Independent and being a swing voter disjoint, i.e. mutually exclusive?
+ No, some voters are Independent and swing voters.
b) Draw a Venn diagram summarizing the variables and their associated probabilities. 

```{r fig.align='center'}
draw.pairwise.venn(area1 = 35, area2 = 23, cross.area = 11, 
                   category = c('% Independent', '% Swing Voter'), 
                   cat.pos = c(3, -1))
```

c) What percent of voters are Independent but not swing voters?
+ 24% are non-swinging Independents.

d) What percent of voters are Independent or swing voters?
+ 47% are Independent and/or swing voters; 36% are Independent or swing voters but not both.

e) What percent of voters are neither Independent nor swing voters?
+ 53% are neither Independent nor swing voters.

f) Is the event that someone is a swing voter independent of the event that someone is a political
Independent?
+ No, the probability of being a swing voter is 23% for all voters, but the probability of being a swing voter given being an independent is 31%. The probabilities would be the same if the two variables were independent. 
