---
title: "Hao-Week7"
author: "Bruce Hao"
date: "October 10, 2016"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(RCurl)
library(XML)
library(jsonlite)
library(dplyr)
setwd("C:/Users/bhao/Google Drive/CUNY/git/DATA607/Week7")
options(width=100)
```

**HTML Table**

```{r}
url_html = getURL('https://raw.githubusercontent.com/haobruce/CUNY/master/DATA607/Week7/Hao-Week7_HTML.html')
books_html = readHTMLTable(url_html, stringsAsFactors = F)[[1]]
books_html
```

**XML**

```{r}
url_xml = getURL('https://raw.githubusercontent.com/haobruce/CUNY/master/DATA607/Week7/Hao-Week7_XML.xml')
data_xml = xmlParse(url_xml)
books_xml = xmlToDataFrame(data_xml)
books_xml

# case with extra parent node
url_xml2 = getURL('https://raw.githubusercontent.com/haobruce/CUNY/master/DATA607/Week7/Hao-Week7_XML2.xml')
data_xml2 = xmlParse(url_xml2)
data_xml2 = xmlRoot(data_xml2)[[1]]
books_xml2 = xmlToDataFrame(data_xml2)
books_xml2
```

**JSON**

```{r}
url_json = getURL('https://raw.githubusercontent.com/haobruce/CUNY/master/DATA607/Week7/Hao-Week7_JSON.json')
books_json = fromJSON(url_json)[[1]]
books_json
```


**Are the three data frames identical?**

* HTML - Using the `readHTMLTable` function from the `XML` package, the output is a List of 1 with the 1 object being the data frame. Additionally, blanks are imported as blanks.
* XML - Using the `xmlParse` and `xmlToDataFrame` functions from the `XML` package, the output is data frame with blanks appearing as <NA>.
* JSON - Using the `fromJSON` function from the `jsonlite` package, the output is a List of 1 with the 1 object being the data frame. Additionally, the blanks are imported as <NA>.


