---
title: "BHao_Assign8"
output: html_document
---

```{r setup, include=FALSE}
setwd("C:/Users/bhao/Google Drive/CUNY/git/DATA604/Week8")
library(xlsx)
library(fitdistrplus)
```

##6.1  

Based on the PP-plots and chi-squared (chisq), Kolmogorov-Smirnov (ks), aic and bic statistics, it appears that the lognormal distribution best fits these data.  

```{r fig.width=10}
df1 = read.xlsx('Problem_Dataset_06_01.xls', 'Sheet1', header = FALSE)

# first plot histogram and CDF 
par(mfrow=c(1,2))
hist(df1$X1)
plot(ecdf(df1$X1))

# based on the histogram, we see that all values are positive with large outliers 
# therefore, we should try to fit exponential, lognormal, gamma and weibull distributions 

# try exponential distribution 
expon1 = fitdist(df1$X1, 'exp', discrete = FALSE)
plot(expon1)

# try log normal distribution 
lnorm1 = fitdist(df1$X1, 'lnorm', discrete = FALSE)
plot(lnorm1)

# try gamma distribution 
gamma1 = fitdist(df1$X1, 'gamma', discrete = FALSE)
plot(gamma1)

# try weibull distribution
weib1 = fitdist(df1$X1, 'weibull', discrete = FALSE)
plot(weib1)

# compare multiple fits 
cbind('exp'=gofstat(expon1), 'lnorm'=gofstat(lnorm1), 'gamma'=gofstat(gamma1), 'weibul'=gofstat(weib1))

# print summary of best fitting distribution
summary(lnorm1)
```

##6.2  

Again, based on the goodness-of-fit statistics and plots, the lognormal distribution seems to fit the data best.  

```{r fig.width=10}
df2 = read.xlsx('Problem_Dataset_06_02.xls', 'Sheet1', header = FALSE)

# first plot histogram and CDF 
par(mfrow=c(1,2))
hist(df2$X1)
plot(ecdf(df2$X1))

# based on the histogram, we see that all values are positive with large outliers 
# therefore, we should try to fit exponential, lognormal, gamma and weibull distributions 

# try exponential distribution 
expon1 = fitdist(df2$X1, 'exp', discrete = FALSE)
plot(expon1)

# try log normal distribution 
lnorm1 = fitdist(df2$X1, 'lnorm', discrete = FALSE)
plot(lnorm1)

# try gamma distribution 
gamma1 = fitdist(df2$X1, 'gamma', discrete = FALSE)
plot(gamma1)

# try weibull distribution
weib1 = fitdist(df2$X1, 'weibull', discrete = FALSE)
plot(weib1)

# compare multiple fits 
cbind('exp'=gofstat(expon1), 'lnorm'=gofstat(lnorm1), 'gamma'=gofstat(gamma1), 'weibul'=gofstat(weib1))

# print summary of best fitting distribution
summary(lnorm1)
```

##6.3  

In this case, the Poisson distribution seemed to fit the data best.  

```{r fig.width=10}
df3 = read.xlsx('Problem_Dataset_06_03.xls', 'Sheet1', header = FALSE)

# first plot histogram and CDF 
par(mfrow=c(1,2))
hist(df2$X1)
plot(ecdf(df2$X1))

# based on the histogram, we see that all values are positive with large outliers 
# also, these data are discrete, we should try discrete distributions  
# therefore, we should try to fit poisson, negative binomial and normal distributions

# try poisson distribution 
pois1 = fitdist(df3$X1, 'pois', discrete = TRUE)
plot(pois1)

# try negative binomial distribution 
nbinom1 = fitdist(df3$X1, 'nbinom', discrete = TRUE)
plot(nbinom1)

# try normal distribution
norm1 = fitdist(df3$X1, 'norm', discrete = TRUE)
plot(norm1)

# compare multiple fits 
cbind('pois'=gofstat(pois1), 'nbinom'=gofstat(nbinom1), 'norm'=gofstat(norm1))

# print summary of best fitting distribution
summary(pois1)
```

##6.4  

\[
  F(x) =
    \begin{cases}
      0 & x \lt a \\
      \frac{x - a}{b - a} & a \le x \le b \\
      1 & x \gt b 
    \end{cases}
\]

\[
  X = a + R (b - a)
\]

##6.5  

\[
  F(x; k, \lambda) = 1 - e^{-(x/\lambda)^k}
\]

\[
  X = -a[ln(1-R)]^{1/\lambda}I_{[0,1]}(R)  
\]

##6.8  

```{r}
set.seed(123)
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)

# create a discrete sampling function 
lbs_sample = function(item) {
  R = runif(1)
  if (item == 'oats') {
    if (R <= 0.05) {lbs = 0.0}
    else if (R <= 0.12) {lbs = 0.5}
    else if (R <= 0.21) {lbs = 1.0}
    else if (R <= 0.32) {lbs = 1.5}
    else if (R <= 0.47) {lbs = 2.0}
    else if (R <= 0.72) {lbs = 3.0}
    else if (R <= 0.82) {lbs = 4.0}
    else if (R <= 0.91) {lbs = 5.0}
    else if (R <= 0.97) {lbs = 7.5}
    else if (R <= 1.00) {lbs = 10.0}
  }
  if (item == 'peas') {
    if (R <= 0.1) {lbs = 0.0}
    else if (R <= 0.3) {lbs = 0.5}
    else if (R <= 0.5) {lbs = 1.0}
    else if (R <= 0.8) {lbs = 1.5}
    else if (R <= 0.9) {lbs = 2.0}
    else if (R <= 1.0) {lbs = 3.0}
  }
  if (item == 'beans') {
    if (R <= 0.2) {lbs = 0.0}
    else if (R <= 0.6) {lbs = 1.0}
    else if (R <= 0.9) {lbs = 3.0}
    else if (R <= 1.0) {lbs = 4.5}
  }
  if (item == 'barley') {
    if (R <= 0.2) {lbs = 0.0}
    else if (R <= 0.6) {lbs = 0.5}
    else if (R <= 0.9) {lbs = 1.0}
    else if (R <= 1.0) {lbs = 3.5}
  }
  return(lbs)
}

# create simulation function for one season
run_sim = function(days = 90) {
  # create data table to house inputs
  tbl_items = data.table(item = c('oats', 'peas', 'beans', 'barley'),
                  cost = c(1.05, 3.17, 1.99, 0.95),
                  price = c(1.29, 3.76, 2.23, 1.65))
  tbl_items = tbl_items[, profit := price - cost]

  # create data frame to house results
  tbl_sim = data.frame(day = rep(0, days), total_revenue = rep(0, days), 
                       total_cost = rep(0, days), total_profit = rep(0, days))

  for (i in 1:days) {
    qty = NULL
    revenue = 0
    cost = 0
    profit = 0
    for (r in 1:nrow(tbl_items)) {
      qty = lbs_sample(tbl_items[r, item])
      revenue = revenue + tbl_items[r, price] * qty
      cost = cost + tbl_items[r, cost] * qty
      profit = profit + tbl_items[r, profit] * qty
    }
    tbl_sim[i,] = c(i, revenue, cost, profit)
  }
  return(data.table(tbl_sim))
}

# create simulation function for x seasons
run_sim_season = function(seasons = 1000, days = 90) {
  tbl_daily = data.frame(day = rep(0, seasons*days), total_revenue = rep(0, seasons*days), 
                         total_cost = rep(0, seasons*days), total_profit = rep(0, seasons*days))
  tbl_season = data.frame(season = rep(0, seasons), total_revenue = rep(0, seasons), 
                          total_cost = rep(0, seasons), total_profit = rep(0, seasons))
  for (i in 1:seasons) {
    ## REPLACE LOGIC BELOW WITH A SIMPLE ARRAY AND THEN RETURN A DATA.TABLE 
    sim = run_sim(days = days)
    tbl_daily[(i*days-days+1):(i*days),] = sim
    tbl_season[i,] = c(season = i, sim[, lapply(.SD, sum), .SDcols = 2:4])
  }
  return(list(tbl_daily = data.table(tbl_daily), tbl_season = data.table(tbl_season)))
}

# run simulation x times and aggregate results
sim_results = run_sim_season(seasons = 100, days = 90)

summary(sim_results$tbl_daily)
summary(sim_results$tbl_season)

sim_results$tbl_daily %>% select(-day) %>%
  gather('metric', 'value') %>%
  mutate('metric' = factor(metric, levels = c('total_revenue', 'total_cost', 'total_profit'))) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_grid(. ~ metric, scales = 'free')

sim_results$tbl_season %>% select(-season) %>%
  gather('metric', 'value') %>%
  mutate('metric' = factor(metric, levels = c('total_revenue', 'total_cost', 'total_profit'))) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_grid(. ~ metric, scales = 'free')
```
