---
title: "Hao-Project4"
output: 
  html_notebook:
    theme: yeti
    toc: true
    toc_float: true
    code_folding: show
---

```{r message=FALSE, warning=TRUE}
library(dbConnect)
library(jsonlite)
library(mongolite)

setwd("~/Google Drive/CUNY/git/DATA607/Project4")
```


```{r echo=FALSE}
username = 'root'
password = 'password'
dbname = 'flights'
host = '127.0.0.1'
```

#Import data from MySQL#  

After loading the flights database into a local MySQL database, I then imported each of the tables into R data frames. Since I'm building the data around the flights table, I only selected those flights originating from EWR since the weather data was really only available for EWR. I also limited it to the first 100 flights since it othewise takes a significant amount of time to format the data and import it into MongoDB.  

```{r}
myDb = dbConnect(MySQL(), user=username, password=password, dbname=dbname, host=host)

flightsQry = dbSendQuery(myDb, 'SELECT * FROM flights WHERE origin = "EWR" LIMIT 100;')
flights = fetch(flightsQry, n = -1)

airportsQry = dbSendQuery(myDb, 'SELECT * FROM airports;')
airports = fetch(airportsQry, n = -1)

airlinesQry = dbSendQuery(myDb, 'SELECT * FROM airlines;')
airlines = fetch(airlinesQry, n = -1)

planesQry = dbSendQuery(myDb, 'SELECT * FROM planes;')
planes = fetch(planesQry, n = -1)

weatherQry = dbSendQuery(myDb, 'SELECT * FROM weather;')
weather = fetch(weatherQry, n = -1)
```

#Convert MySQL tables into nested data frames#  

Then, in order to nest the airport, airline, plane and weather details, I looped through the flights data frame and nested the aforementioned details as embedded data frames. To verify that this nesting process would produce the JSON format I wanted, I used the `prettify` and `toJSON` functions to confirm.  

I initially tried and failed to apply a vectorized approach to this procedure. Furthermore, I also tried and failed to use the `apply` family of functions to accomplish this. Any help would be much appreciated regarding this, as I seem to often struggle to find more efficient approaches to these types of challenges.  

```{r}
for (i in 1:nrow(flights)) {
  flights$origin_detail[[i]] = airports[airports$faa == flights$origin[i],]
  flights$dest_detail[[i]] = airports[airports$faa == flights$dest[i],]
  flights$airline_detail[[i]] = airlines[airlines$carrier == flights$carrier[i],]
  flights$plane_detail[[i]] = planes[planes$tailnum == flights$tailnum[i],]
  # check if weather row is NA
  weather_row = weather[weather$origin == flights$origin[i] & 
                                   weather$year == flights$year[i] &
                                   weather$month == flights$month[i] &
                                   weather$day == flights$day[i] &
                                   weather$hour == flights$hour[i]
                                   , ][1,]
  if (!is.na(weather_row[1])) {
    flights$weather_detail[[i]] = weather[weather$origin == flights$origin[i] & 
                                   weather$year == flights$year[i] &
                                   weather$month == flights$month[i] &
                                   weather$day == flights$day[i] &
                                   weather$hour == flights$hour[i]
                                   , ][1,]
  } 
}

# check JSON validity
prettify(toJSON(flights[1:3,]))
```

#Import JSON file into MongoDB#  

Having confirmed that the JSON output was valid, I then researched potential MongoDB packages for R and settled on mongolite as it builds on jsonlite, which I found to be more user-friendly than the alternatives. I also installed MongoDB locally and started a mongod service in order to import data and interact with them. Lastly, I run a few simple queries and then finally drop the collection.  

```{r}
m = mongo(collection = 'flights')
m$insert(flights)
# check record counts
m$count()

# perform query and retrieve data
m$find( '{ "carrier": "UA" }' )
m$find( ' { "plane_detail.manufacturer" : "BOEING" } ' )

# list distinct values
m$distinct("plane_detail.manufacturer")

# remove the collection
m$drop()
```
