---
title: "BHao_Final"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
setwd("C:/Users/bhao/Google Drive/CUNY/git/DATA605/Final")
```

```{r}
library(dplyr)
library(ggplot2)
library(gridExtra)

kaggle = read.csv("C:/Users/bhao/Google Drive/CUNY/git/DATA605/Final/train.csv")
#str(kaggle)
```

###Probability  

Since P(X|Y) != P(X)P(Y), splitting the data this way does not make them independent. This is also confirmed by the chi-square test, which resulted in a very small p-value.  

```{r}
library(MASS)

df = kaggle %>% dplyr::select(SalePrice, LotArea)
X = df$LotArea
Y = df$SalePrice

# set x = min of 4th quartile and y = min of 2nd quartile 
x = quantile(df$LotArea, 0.75)
y = quantile(df$SalePrice, 0.25)

#a. P(X>x | Y>y) = 31.23% probability that lot area is greater than 0.75 quantile when 
#sale price is greater than 0.25 quantile 
df %>% filter(LotArea > x & SalePrice > y) %>% summarise(n = n()) / 
df %>% filter(SalePrice > y) %>% summarise(n = n())

#b. P(X>x, Y>y) = 23.42% probability that lot area is greater than 0.75 quantile AND
#sale price is greater than 0.25 quantile
df %>% filter(LotArea > x & SalePrice > y) %>% summarise(n = n()) / 
nrow(df)

#c. P(X<x | Y>y) = 68.77% probability that lot area is less than 0.75 quantile when
#sale price is greater than 0.25 quantile
df %>% filter(LotArea < x & SalePrice > y) %>% summarise(n = n()) / 
df %>% filter(SalePrice > y) %>% summarise(n = n())

#chi-square test
tbl = table(df %>% mutate(topQtrX = if_else(LotArea > x, 'TopQtr', 'Bottom3Qtr'),
              botQtrY = if_else(SalePrice > y, 'Top3Qtr', 'BottomQtr')) %>%
  dplyr::select(topQtrX, botQtrY))

chisq.test(tbl)
```

###Descriptive and Inferential Statistics  

```{r}
summary(df)
p1 = df %>% ggplot(aes(x = factor(0), y = SalePrice)) + geom_boxplot()
p2 = df %>% ggplot(aes(x = factor(0), y = LotArea)) + geom_boxplot()
grid.arrange(p1, p2, ncol = 2)

df %>% ggplot(aes(x = LotArea, y = SalePrice)) + geom_point()

#box-cox transformation
bc = boxcox(SalePrice ~ LotArea, data = df)
cor(bc$x, bc$y, method = 'spearman')
cor(df$LotArea, df$SalePrice)
```

###Linear Algebra and Correlation  

```{r}
cor_mat = cor(df)
prec_mat = solve(cor_mat)  # invert correlation matrix
cor_mat %*% prec_mat  # results in identity matrix
prec_mat %*% cor_mat  # results in identity matrix
```

###Calculus-Based Probability & Statistics  

I fit a lognormal distribution to the lot area data; the actual data exhibits much more skew wit a fatter tail.  

```{r}
# fit lognormal distribution to data 
fit = fitdistr(df$LotArea, densfun = 'lognormal')
meanLog = fit$estimate[1]
sdLog = fit$estimate[2]

samp = rlnorm(1000, meanLog, sdLog)
typeof(samp)
p1 = df %>% ggplot(aes(x = LotArea)) + geom_histogram() + xlim(c(0, 50000))
p2 = qplot(samp, geom = 'histogram')
grid.arrange(p1, p2, ncol = 2)
```

###Modeling  

```{r}
library(caret)

set.seed(123)
myControl = trainControl(method = 'cv', number = 5, verboseIter = TRUE)

#try glmnet model 
clean = imputeMissings::impute(kaggle, method = 'median/mode')  # impute medians for missing data 

glmnet_model = train(SalePrice ~ ., data = clean, method = 'glmnet', trControl = myControl,
                     preProcess = c('center', 'scale'))

#try random forest model 
rf_model = train(SalePrice ~ ., data = clean, method = 'ranger', trControl = myControl,
                     preProcess = c('center', 'scale'))

#compare models 
model_list = list(glmnet = glmnet_model, rf = rf_model)

#collect resamples from the CV folds
resamps = resamples(model_list)
dotplot(resamps, metric = 'RMSE')

#load test data 
test = read.csv("C:/Users/bhao/Google Drive/CUNY/git/DATA605/Final/test.csv")
test_clean = imputeMissings::impute(test, method = 'median/mode')  # impute medians for missing data 
test_pred = predict(rf_model, newdata = test_clean)
test_pred = cbind(test, test_pred) %>% mutate(SalePrice = test_pred) %>% dplyr::select(Id, SalePrice)
write.csv(test_pred, "C:/Users/bhao/Google Drive/CUNY/git/DATA605/Final/rf_pred.csv")
```


