---
title: "Hao-Week5"
author: "Bruce Hao"
date: "September 23, 2016"
output: github_document
---

```{r message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggthemes)
```

**1. Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You’re encouraged to use a “wide” structure similar to how the information appears above, so that you can practice tidying and transformations as described below.**

Create data frame housing airline status information. Write data frame to csv file. Read csv file back to data frame. I commented it out to prevent the path name from causing an error.

```{r}
airlines = data.frame(
  airline = c('Alaska', 'Alaska', 'AM West', 'AM West'),
  status = c('on time', 'delayed', 'on time', 'delayed'),
  LosAngeles = c(497, 62, 694, 117),
  Phoenix = c(221, 12, 4840, 415),
  SanDiego = c(212, 20, 383, 65),
  SanFrancisco = c(503, 102, 320, 129),
  Seattle = c(1841, 305, 201, 61)
)

#write.csv(airlines, file = '/Users/brucehao/Desktop/airlines.csv', row.names = F)
#airlines = read.csv('/Users/brucehao/Desktop/airlines.csv')

airlines
```


**2. Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.**

I've broken the tidying exercise up step by step to help visualize the process. In part 3 below, the intermediate steps are hidden.

1) gather all flights by adding an 'airport' key and 'flights' value; 

```{r}
airlines2 = airlines %>% 
  gather(airport, flights, -(1:2)) 
airlines2
```

2) we then spread the status column into two new columns 'delayed' and 'on time'; 

```{r}
airlines3 = airlines2 %>% 
  spread(status, flights) 
airlines3
```

3) then, using the mutate function we create a 'total' flights column; 

```{r}
airlines4 = airlines3 %>%
  mutate('total' = delayed + `on time`) 
airlines4
```

and 4) finally we group by airline and summarize the total on time flights divided by the total flights.
```{r}
airlines5 = airlines4 %>%
  group_by(airline) %>% 
  summarise(OnTimePct = sum(`on time`) / sum(total))
airlines5
```


**3. Perform analysis to compare the arrival delays for the two airlines.**

First, let's evaluate each airline's overall performance in terms of ratio of total on time arrivals to total flights. 

Based on this particular analysis, it appears that AM West has a slightly better on time ratio than does Alaska.

```{r fig.height=3}
airlines %>% 
  gather(airport, flights, -(1:2)) %>% 
  spread(status, flights) %>%
  mutate('total' = delayed + `on time`) %>%
  group_by(airline) %>% 
  summarise(OnTimePct = sum(`on time`) / sum(total)) %>%
  ggplot(aes(x = airline, y = OnTimePct, fill = airline)) + 
    geom_bar(stat = 'identity', width = 0.5) + 
    geom_text(aes(label = format(OnTimePct, digits = 2)), vjust = -0.25) + 
    scale_y_continuous(limits = c(0, 1)) +
    scale_fill_economist() +
    theme_minimal()
```

Next, let's evaluate each airlines performance on an airport by airport basis. In this case, we follow the same steps except that we group by both airline and airport. Additionally, we plot airport on the x-axis instead of airline and use color coding to indicate the airline. 

Based on this analysis, it's now clear that Alaska actually has a better on time ratio than AM West at each of the five airports. 

```{r fig.height=3}
airlines %>% 
  gather(airport, flights, -(1:2)) %>% 
  spread(status, flights) %>%
  mutate('total' = delayed + `on time`) %>%
  group_by(airline, airport) %>% 
  summarise(OnTimePct = sum(`on time`) / sum(total)) %>%
  ggplot(aes(x = airport, y = OnTimePct, fill = airline)) + 
    geom_bar(stat = 'identity', width = 0.75, position = 'dodge') + 
    geom_text(aes(label = format(OnTimePct, digits = 2)), position = position_dodge(width=0.75), vjust = -0.25) +
    scale_y_continuous(limits = c(0, 1)) +
    scale_fill_economist() +
    theme_minimal()
```

The chart below tries to explain this apparent contradiction. The width of the bars below represents the total number of flights for each airline at each airport. AM West obviously makes a lot of flights to/from Phoenix, where on time performance is generally better than at the other airports for both airlines. Taken together, these factors combine to 'pull up' AM West's overall on time performance and hides the fact that it actually performs worse than does Alaska at each and every airport, including Phoenix. 

```{r fig.height=3, warning=FALSE}
airlines %>% 
  gather(airport, flights, -(1:2)) %>% 
  spread(status, flights) %>%
  mutate('total' = delayed + `on time`) %>%
  group_by(airline, airport) %>% 
  summarise(OnTimePct = sum(`on time`) / sum(total), Total = sum(total)) %>%
  ggplot(aes(x = airport, y = OnTimePct, fill = airline, width = Total/5000)) + 
    geom_bar(stat = 'identity') + 
    geom_text(aes(label = format(OnTimePct, digits = 2), color = airline), position = position_dodge(width=0.75), vjust = -0.25) +
    scale_y_continuous(limits = c(0, 1)) +
    scale_fill_economist() +
    scale_color_economist() +
    theme_minimal()
```


**4. Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative descriptions of your data cleanup work, analysis, and conclusions.**

