---
title: "BHao_Assign11"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
setwd("C:/Users/bhao/Google Drive/CUNY/git/DATA605/HW11")
```

* Fitting max heart rate to age using lm results in the following equation: max heart rate = 209.3416 - 0.7744 * age  
* The effect of age on max heart rate is significant  
* The significant level is at the 0.001 level  

```{r}
age = c(18,23,25,35,65,54,34,56,72,19,23,42,18,39,37)
maxHR = c(202,186,187,180,156,169,174,172,156,199,193,174,198,183,178)

age_hr = lm(maxHR ~ age)
summary(age_hr)
plot(maxHR ~ age)
abline(age_hr$coefficients[1], age_hr$coefficients[2])
```

* Fitting a multiple regression the auto data set results in the equation below: mpg = 45.2511 - 0.0060 * disp - 0.0436 * hp - 0.0053 * weight - 0.0231 * accel  
* hp and weight have significant impacts  
* hp and weight have 0.001 and 0.01 significance levels, respectively  
* the standard errors for disp, hp, weight and accel are shown below or 0.0067, 0.0166, 0.0008 and 0.1256, respectively  

```{r}
mpg = read.table('auto-mpg.data')
names(mpg) = c('disp', 'hp', 'weight', 'accel', 'mpg')
str(mpg)

lm_model = lm(mpg ~ ., data = mpg)
summary(lm_model)
```

