---
title: "BHao_HW5"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
setwd("C:/Users/bhao/Google Drive/CUNY/git/DATA621/HW5")
```

##DATA EXPLORATION  

* There are many NAs scattered throughout the data set; we'll need to decide on how best to handle them  
* First, let's take a look at the distribution of the TARGET variable  
* Next, we only look at those observations where data is missing and review the TARGET variable distribution  
* It appears that the distributions are mostly similar except in the case of the STARS variable, where there is a much greater proportion of 0 cases as compared to the other variables  
* As such, we may be able to impute medians for the other variables, but in the case of the STARS variable, we'll set the missing data equal to 0 since there are no other zeros within this variable  

```{r warning=FALSE, message=FALSE, fig.height = 4, fig.width= 10}
library(dplyr)
library(ggplot2)
library(gridExtra)

wine = read.csv('wine-training-data.csv', stringsAsFactors = TRUE)
# drop index column 
wine = subset(wine, select = -c(ï..INDEX))
str(wine)
summary(wine)

p1 = wine %>% ggplot(aes(x = TARGET)) + geom_histogram() + ggtitle('All Data')
p2 = wine %>% filter(is.na(ResidualSugar)) %>% ggplot(aes(x = TARGET)) + geom_histogram() + 
  ggtitle('ResidualSugar == NA')
p3 = wine %>% filter(is.na(Chlorides)) %>% ggplot(aes(x = TARGET)) + geom_histogram() +
  ggtitle('Chlorides == NA')
p4 = wine %>% filter(is.na(FreeSulfurDioxide)) %>% ggplot(aes(x = TARGET)) + geom_histogram() +
  ggtitle('FreeSulfurDioxide == NA')
p5 = wine %>% filter(is.na(TotalSulfurDioxide)) %>% ggplot(aes(x = TARGET)) + geom_histogram() +
  ggtitle('TotalSulfurDioxide == NA')
p6 = wine %>% filter(is.na(pH)) %>% ggplot(aes(x = TARGET)) + geom_histogram() +
  ggtitle('pH == NA')
p7 = wine %>% filter(is.na(Sulphates)) %>% ggplot(aes(x = TARGET)) + geom_histogram() +
  ggtitle('Sulphase == NA')
p8 = wine %>% filter(is.na(Alcohol)) %>% ggplot(aes(x = TARGET)) + geom_histogram() +
  ggtitle('Alcohol == NA')
p9 = wine %>% filter(is.na(STARS)) %>% ggplot(aes(x = TARGET)) + geom_histogram() +
  ggtitle('STARS == NA')

grid.arrange(p1, p2, p3, ncol = 3)
grid.arrange(p4, p5, p6, ncol = 3)
grid.arrange(p7, p8, p9, ncol = 3)
```

##DATA PREPARATION  

* We'll first impute medians for the non-STARS missing data  
* We'll then set the NAs in STARS to 0  
* Lastly, we'll add a flag to indicate STARS == 0  

```{r}
# median imputation
impute_median = function(x) replace(x, is.na(x), median(x, na.rm = TRUE))
wine$ResidualSugar = impute_median(wine$ResidualSugar)
wine$Chlorides = impute_median(wine$Chlorides)
wine$FreeSulfurDioxide = impute_median(wine$FreeSulfurDioxide)
wine$TotalSulfurDioxide = impute_median(wine$TotalSulfurDioxide)
wine$pH = impute_median(wine$pH)
wine$Sulphates = impute_median(wine$Sulphates)
wine$Alcohol = impute_median(wine$Alcohol)
wine[is.na(wine$STARS), 'STARS'] = 0
wine$STARS_FLAG = if_else(wine$STARS == 0, 'NoStars', 'Stars')
summary(wine)
```

##BUILD MODELS  

* We'll first build a count regression model to estimate the number of cases sold  
* In terms of setup, we are using 10-fold cross validation to measure out-of-sample performance and are using the same folds for each model to ensure comparable results    
* We then start by including all variables and then remove statistically insignificant ones at the 5% level until all remaining are significant  
* We then tried a glmnet model which combines lasso and ridge regression; given that it penalizes large magnitude and the number of non-zero coefficients, it can be used for variable selection  
* Lastly, we fit a random forest model just for fun  
* Based on the RMSE dot plot, there does not appear to be much improvement as the model is simplified; still, for the sake of parsimony, we'll use the simplest version as our final model  
* Note how well the rf model performed without manual tuning compared to the glmnet model which performed the poorest

```{r message=FALSE, warning=FALSE}
library(caret)
library(caretEnsemble)

set.seed(123)
# use cross validation to compare out-of-sample ROC for all models  
# use the same folds for each model to ensure comparable results 
myFolds = createFolds(wine$TARGET, k = 10)  

# used instead of method = 'cv', number = 10
myControl = trainControl(verboseIter = FALSE, savePredictions = TRUE, index = myFolds)  

# model using glm model
model_glm_full = train(TARGET ~ ., data = wine, method = 'glm', family = 'poisson',
                  preProcess = c('center', 'scale'), trControl = myControl)
summary(model_glm_full)

# let's drop any statistically insignificant variables at 5% 
model_glm_sig1 = train(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide + TotalSulfurDioxide +
                       pH + Sulphates + LabelAppeal + AcidIndex + STARS + STARS_FLAG, 
                       data = wine, method = 'glm', family = 'poisson',
                       preProcess = c('center', 'scale'), trControl = myControl)
summary(model_glm_sig1)

# let's again drop any additional statistically insigificant variables at 10% 
model_glm_sig2 = train(TARGET ~ VolatileAcidity + FreeSulfurDioxide + TotalSulfurDioxide +
                       LabelAppeal + AcidIndex + STARS + STARS_FLAG, 
                       data = wine,
                       method = 'glm', family = 'poisson',
                       preProcess = c('center', 'scale'), trControl = myControl)
summary(model_glm_sig2)

# let's try a glmnet model that combines ridge vs. lasso regression 
# since it penalizes either or both magnitude and number of non-zero coefficients, it can be used for variable selection
model_glmnet = train(TARGET ~ ., data = wine, method = 'glmnet', family = 'poisson',
                     preProcess = c('center', 'scale'), trControl = myControl)
coef(model_glmnet$finalModel, s = model_glmnet$finalModel$tuneValue$lambda)

# let's also model using random forest just for fun 
model_rf = train(TARGET ~ ., data = wine, method = 'ranger', 
                 trControl = myControl)

# compare models 
model_list = list(glm_full = model_glm_full, glm_sig1 = model_glm_sig1, glm_sig2 = model_glm_sig2, 
                  glmnet = model_glmnet, rf = model_rf)

# collect resamples from the CV folds
resamps = resamples(model_list)
summary(resamps)
dotplot(resamps, metric = 'RMSE')
```

##SELECT MODEL  

* The final models were selected because they performed the best, are very simple and are highly intuitive  
* Since the variables were centered and scaled, we can interpret the coefficients on a more apples-to-apples basis:  
    - STARS had the highest explanatory effect, with the more stars the more cases sold  
    - Label appeal was the second most explanatory again with a higher appeal associated with more cases sold  
    - AcidIndex was the third with less acid associated with more cases sold  
* After the final models were selected, we then re-fit the models to the entire data set (i.e. no cross validation) to ensure that we maximize use of all the available data   
* The final logistic regression model is then used to predict the classes and probabilities  
* Finally, the final linear regression model is used to predict the cost of damage for only those predicted accidents  

```{r}
final_model = train(TARGET ~ VolatileAcidity + FreeSulfurDioxide + TotalSulfurDioxide +
                     LabelAppeal + AcidIndex + STARS + STARS_FLAG,
                     data = wine,
                     method = 'glm', family = 'poisson',
                     preProcess = c('center', 'scale'), 
                     trControl = trainControl(verboseIter = FALSE))
summary(final_model)

# import and cleanse test data 
wine_test = read.csv('wine-evaluation-data.csv', stringsAsFactors = TRUE)
wine_test$ResidualSugar = impute_median(wine_test$ResidualSugar)
wine_test$Chlorides = impute_median(wine_test$Chlorides)
wine_test$FreeSulfurDioxide = impute_median(wine_test$FreeSulfurDioxide)
wine_test$TotalSulfurDioxide = impute_median(wine_test$TotalSulfurDioxide)
wine_test$pH = impute_median(wine_test$pH)
wine_test$Sulphates = impute_median(wine_test$Sulphates)
wine_test$Alcohol = impute_median(wine_test$Alcohol)
wine_test[is.na(wine_test$STARS), 'STARS'] = 0
wine_test$STARS_FLAG = if_else(wine_test$STARS == 0, 'NoStars', 'Stars')
summary(wine_test)

# predict cases sold
pred = predict(final_model, newdata = wine_test)
wine_test$TARGET = pred

write.csv(wine_test, 'wine-evaluation-prediction.csv')

# check the distribution of predicted cases sold 
wine_test %>% ggplot(aes(x = TARGET)) + geom_histogram()
```


