---
title: "BHao_Assign8"
output: html_document
---

```{r setup, include=FALSE}
setwd("C:/Users/bhao/Google Drive/CUNY/git/DATA605/HW8")
```

##Problem Set 1  

+/-B = takes bus (does not take train); does not take bus (takes train)   
+/-L = late; not late  

Solve for P(+B | -L) = P(+B, -L) / P(-L)

Solve numerator:  

P(+B, -L) = 1/3 * 0.5  

Solve denominator:  

P(-L) = $\sum$$_B$ P(-L)  
P(-L) = $\sum$$_B$ P(-L | B)  
[case: +B, -L]: + P(+B) P(-L | +B)  
[case: -B, -L]: + P(-B) P(-L | -B)  

Given that she was on time today, the probability that she took the bus to work today was 21.74%.  

```{r}
numer = 1/3 * 0.5  
denom = 1/3 * 0.5 + 2/3 * 0.9
numer / denom
```


##Problem Set 2  

*What happens to the probability of Dificulty of Course when you present the evidence that the received recommendation letter was good?*  

+/- D = not difficult; yes difficult  
+/- I = low intelligence; high intelligence  
+/- G = low grade; high grade  
+/- S = low SAT; high SAT  
+/- L = bad letter of rec; good letter of rec  

Solve for P(+D | +L) = P(+D, +L) / P(+L)  

Solve numerator:  

P(+D, +L) = $\sum$$_G$ $\sum$$_I$ P(+D, +L)  
P(+D, +L) = $\sum$$_G$ $\sum$$_I$ P(+D) P(I) P(G | +D, I) P(+L | G)  
P(+D, +L) =  
[case: +G, +I]: + P(+D) P(+I) P(+G | +D, +I) P(+L | +G)  
[case: +G, -I]: + P(+D) P(-I) P(+G | +D, -I) P(+L | +G)  
[case: -G, +I]: + P(+D) P(+I) P(-G | +D, +I) P(+L | -G)  
[case: -G, -I]: + P(+D) P(-I) P(-G | +D, -I) P(+L | -G)  

```{r}
numer1 = 0.7 * 0.2 * 0.9 * 0.95  
numer2 = 0.7 * 0.8 * 0.2 * 0.95
numer3 = 0.7 * 0.2 * 0.1 * 0.1
numer4 = 0.7 * 0.8 * 0.8 * 0.1
numerator = numer1 + numer2 + numer3 + numer4
```

Solve denominator:  

P(+L) = $\sum$$_D$ $\sum$$_I$ $\sum$$_G$ P(+L)  
P(+L) = $\sum$$_D$ $\sum$$_I$ $\sum$$_G$ P(D) P(I) P(G | D, I) P(+L | G)  
P(+L) = 
[case: +D, +I, +G]: + P(+D) P(+I) P(+G | +D, +I) P(+L | +G)  
[case: +D, -I, +G]: + P(+D) P(-I) P(+G | +D, -I) P(+L | +G)  
[case: +D, +I, -G]: + P(+D) P(+I) P(-G | +D, +I) P(+L | -G)  
[case: +D, -I, -G]: + P(+D) P(-I) P(-G | +D, -I) P(+L | -G)  
[case: -D, +I, +G]: + P(-D) P(+I) P(+G | -D, +I) P(+L | +G)  
[case: -D, -I, +G]: + P(-D) P(-I) P(+G | -D, -I) P(+L | +G)  
[case: -D, +I, -G]: + P(-D) P(+I) P(-G | -D, +I) P(+L | -G)  
[case: -D, -I, -G]: + P(-D) P(-I) P(-G | -D, -I) P(+L | -G)  

```{r}
denom1 = 0.7 * 0.2 * 0.9 * 0.95
denom2 = 0.7 * 0.8 * 0.2 * 0.95
denom3 = 0.7 * 0.2 * 0.1 * 0.1
denom4 = 0.7 * 0.8 * 0.8 * 0.1
denom5 = 0.3 * 0.2 * 0.99 * 0.95
denom6 = 0.3 * 0.8 * 0.4 * 0.95
denom7 = 0.3 * 0.2 * 0.01 * 0.1
denom8 = 0.3 * 0.8 * 0.6 * 0.1
denominator = denom1 + denom2 + denom3 + denom4 + denom5 + denom6 + denom7 + denom8
```

Given that the letter of recommendation was good, the probability that the course was difficult was 62.69%.  

```{r}
numerator / denominator
```

*Repeat using gRain package in R*  

```{r}
library(gRain)

lh = c('low', 'high')
D = cptable(~Difficulty, values = c(0.3, 0.7), levels = lh)
I = cptable(~Intelligence, values = c(0.8, 0.2), levels = lh)
G.DI = cptable(~Grade|Intelligence:Difficulty, values = c(0.6, 0.4, 0.01, 0.99, 0.8, 0.2, 0.1, 0.9), levels = lh)
L.G = cptable(~Letter|Grade, values = c(0.9, 0.1, 0.05, 0.95), levels = lh)
S.I = cptable(~SAT|Intelligence, values = c(0.9, 0.1, 0.2, 0.8), levels = lh)

plist = compileCPT(list(D, I, G.DI, L.G, S.I))
plist
plist$Grade  # check order of conditional probabilities
plist$Letter  # check order of conditional probabilities  
```

```{r}
net1 = grain(plist)
querygrain(net1, nodes = c('Difficulty', 'Intelligence', 'SAT', 'Grade', 'Letter'), type = 'marginal')
```

```{r}
# query joint distribution  
querygrain(net1, nodes = c('Letter', 'Grade'), type = 'joint')
```

```{r}
# add evidence  
net12 = setEvidence(net1, evidence = list(Letter='high'))

# probability of observing this evidence under the model  
pEvidence(net12)

# query network again given new evidence  
querygrain(net12, nodes = c('Difficulty'), type = 'marginal')  
```

**gRain results tie with manual calculations above!**  

*In addition, now present the evidence that both SAT scores were good and the letter of recommendation was good, What is the probability of the Difficulty of Course now?*  

Manually, we need to update the probabilities for the Intelligence node based on the new SAT information.  

P(+I | +S) = P(+I, +S) / P(+S)  

P(+I, +S) = P(+I) P(+S | +I) 
P(+S) = $\sum$$_I$ P(+S)  
P(+S) = $\sum$$_I$ P(I) P(+S | I)  
[case: +I] + P(+I) P(+S | +I)  
[case: -I] + P(-I) P(+S | -I)  

Updated probabilities for +/-I are 2/3 and 1/3 respectively.  

```{r}
numerator = 0.2 * 0.8  
denominator = (0.2 * 0.8) + (0.8 * 0.1)  
numerator / denominator  
```

Solve for P(+D | +L) = P(+D, +L) / P(+L)  

Solve numerator:  

P(+D, +L) = $\sum$$_G$ $\sum$$_I$ P(+D, +L)  
P(+D, +L) = $\sum$$_G$ $\sum$$_I$ P(+D) P(I) P(G | +D, I) P(+L | G)  
P(+D, +L) =  
[case: +G, +I]: + P(+D) P(+I) P(+G | +D, +I) P(+L | +G)  
[case: +G, -I]: + P(+D) P(-I) P(+G | +D, -I) P(+L | +G)  
[case: -G, +I]: + P(+D) P(+I) P(-G | +D, +I) P(+L | -G)  
[case: -G, -I]: + P(+D) P(-I) P(-G | +D, -I) P(+L | -G)  

```{r}
numer1 = 0.7 * 2/3 * 0.9 * 0.95  
numer2 = 0.7 * 1/3 * 0.2 * 0.95
numer3 = 0.7 * 2/3 * 0.1 * 0.1
numer4 = 0.7 * 1/3 * 0.8 * 0.1
numerator = numer1 + numer2 + numer3 + numer4
```

Solve denominator:  

P(+L) = $\sum$$_D$ $\sum$$_I$ $\sum$$_G$ P(+L)  
P(+L) = $\sum$$_D$ $\sum$$_I$ $\sum$$_G$ P(D) P(I) P(G | D, I) P(+L | G)  
P(+L) = 
[case: +D, +I, +G]: + P(+D) P(+I) P(+G | +D, +I) P(+L | +G)  
[case: +D, -I, +G]: + P(+D) P(-I) P(+G | +D, -I) P(+L | +G)  
[case: +D, +I, -G]: + P(+D) P(+I) P(-G | +D, +I) P(+L | -G)  
[case: +D, -I, -G]: + P(+D) P(-I) P(-G | +D, -I) P(+L | -G)  
[case: -D, +I, +G]: + P(-D) P(+I) P(+G | -D, +I) P(+L | +G)  
[case: -D, -I, +G]: + P(-D) P(-I) P(+G | -D, -I) P(+L | +G)  
[case: -D, +I, -G]: + P(-D) P(+I) P(-G | -D, +I) P(+L | -G)  
[case: -D, -I, -G]: + P(-D) P(-I) P(-G | -D, -I) P(+L | -G)  

```{r}
denom1 = 0.7 * 2/3 * 0.9 * 0.95
denom2 = 0.7 * 1/3 * 0.2 * 0.95
denom3 = 0.7 * 2/3 * 0.1 * 0.1
denom4 = 0.7 * 1/3 * 0.8 * 0.1
denom5 = 0.3 * 2/3 * 0.99 * 0.95
denom6 = 0.3 * 1/3 * 0.4 * 0.95
denom7 = 0.3 * 2/3 * 0.01 * 0.1
denom8 = 0.3 * 1/3 * 0.6 * 0.1
denominator = denom1 + denom2 + denom3 + denom4 + denom5 + denom6 + denom7 + denom8
```

Given a good letter of recommendation and high SAT score, the probability that the course was difficult was 66.77%.  

```{r}
numerator / denominator
```

*Repeat using gRain package in R*  

```{r}
# add evidence  
net13 = setEvidence(net1, evidence = list(Letter='high', SAT='high')) 

# query neetwork again given new evidence
querygrain(net13, nodes = c('Difficulty'), type = 'marginal')  
```

**gRain results tie with manual calculations above!**  

