---
title: "Hao-6"
output: html_notebook
---

```{r warning=FALSE, message=FALSE}
library(IS606)
library(dplyr)
library(tidyr)
setwd("C:/Users/bhao/Google Drive/CUNY/git/DATA606/Lab6")
load('more/atheism.RData')
```


**Exercise1: In the first paragraph, several key findings are reported. Do these percentages appear to be sample statistics (derived from the data sample) or population parameters?**  

These appear to be sample statistics. To be population parameters, the survey would have had to ask every last person within each country covered.  


**Exercise2: The title of the report is "Global Index of Religiosity and Atheism". To generalize the report's findings to the global human population, what must we assume about the sampling method? Does that seem like a reasonable assumption?**  

We would have to assume that the sample was representative of the global population. The survey covered 57 countries out of the ~200 or so that exist. Of course, regions within countries can be significantly different from one another as well. They survey also does not say anything about age, which can also be a relevant factor. Overall, it's very hard to say if this survey represents the entire human population well, but given the scope of the survey, it's probably as good as one can expect.  


**Exercise3: What does each row of Table 6 correspond to? What does each row of atheism correspond to?**  

Row 6 in the atheism data frame corresponds to an individual respondent in Afghanistan.

```{r}
atheism[6,]
```


**Exercise4: Using the command below, create a new dataframe called us12 that contains only the rows in atheism associated with respondents to the 2012 survey from the United States. Next, calculate the proportion of atheist responses. Does it agree with the percentage in Table 6? If not, why?**

Table 6 actually shows 5% for 'convinced atheists' for the United States, and this matches data.  

```{r}
us12 <- subset(atheism, nationality == "United States" & year == "2012")
us12 %>% group_by(response) %>%
  summarise(n = n()) %>%
  mutate(freq = round(n / sum(n), 2))
```


**Exercise5: Write out the conditions for inference to construct a 95% confidence interval for the proportion of atheists in the United States in 2012. Are you confident all conditions are met?**  

* The individual observations must be independent. A random sample from less than 10% of the poulation ensures independence. This condition is met.  
* Size and skew. The more skewed the data, the larger the sample size need be in order to satisfy the conditions for inference. This condition is met as well.  
* Success-failure condition. The sample sized must also be sufficiently large, which is checked using the success-failure condition. Both proportions well exceed n=10, so this condition is also met.  

```{r}
inference(us12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")
```

```{r}
# manual calculation of standard error of proportion
p = us12 %>% group_by(response) %>%
  summarise(n = n()) %>%
  mutate(freq = round(n / sum(n), 2))
p = p[1, 3]
n = us12 %>% summarise(n = n())
n = n[1, 1]
se = sqrt(p * (1 - p) / n)
se
```


**Exercise6: Based on the R output, what is the margin of error for the estimate of the proportion of the proportion of atheists in US in 2012?**  

The margin of error is more like +/- 1.35%

```{r}
se * qnorm(0.025)
```

**Exercise7: Using the inference function, calculate confidence intervals for the proportion of atheists in 2012 in two other countries of your choice, and report the associated margins of error. Be sure to note whether the conditions for inference are met. It may be helpful to create new data sets for each of the two countries first, and then use these data sets in the inference function to construct the confidence intervals.**  

```{r}

inference(us12$response, est = "proportion", type = "ci", method = "theoretical", 
          success = "atheist")

```

