---
title: "Hao-Project1"
author: "Bruce Hao"
date: "September 10, 2016"
output: html_document
---

```{r warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RCurl)
library(readr)
library(stringr)
library(dplyr)
library(tidyr)
library(DT)
```

Download and store raw data from github repository. `read.csv` was producing an error until I added the `text=` in front of the url. I don't completely understand this behavior.

```{r warning=FALSE, message=FALSE}
# COMMENTED OUT CODE READR
# readLines reads txt file in as string (while maintaining rows) instead of list that read.csv does
#raw = readLines('https://raw.githubusercontent.com/haobruce/CUNY/master/DATA607/Project1/tournamentinfo.txt')
#raw = str_replace_all(raw, '-', '')

# read.delim splits string on some delimiter into a list of multiple columns and rows
#input = read.delim(textConnection(raw), header = F, sep = '|', stringsAsFactors = F)
# drop the last column
#input = input[,-11]

url = getURL('https://raw.githubusercontent.com/haobruce/CUNY/master/DATA607/Project1/tournamentinfo.txt')
chess2 = unlist(read.csv(text=url, stringsAsFactors = F))  # convert output to string
chess2 = str_replace_all(chess2, '-', '')
chess2 = str_split(chess2, '\\|', simplify = T)
chess2 = trimws(chess2)
input = as.data.frame(chess2, stringsAsFactors = F)
input = input[input[,1] != '', ]  # delete empty rows
input = input[, -11]  # drop last column

# first two rows make up the column names
header = paste(str_trim(as.character(input[1,])), str_trim(as.character(input[2,])))
header = make.names(str_sub(header, 1, 11))  # limit header name to 11 characters for Player.Name column
names(input) = header
# drop first two rows
input = input[-1:-2,]

# create vector to house player ID and state
player_id = seq(1, nrow(input), 2)
state_rank_id = seq(2, nrow(input), 2)
state_rank = select(input[state_rank_id, ], State = Pair.Num, Rank = Player.Name)
state_rank$Rank = str_extract(state_rank$Rank, 'R:\\s\\d+')
state_rank$Rank = str_extract(state_rank$Rank, '\\d+')

# apply player_id vector to input to isolate player rows; combine with states
output = input[player_id, ]
output = bind_cols(output, as.data.frame(state_rank))

# calculate average opponent rankings
output_tidy = output %>%
  gather('Round', 'OppNum', 4:10) %>%
  mutate(OppNum = as.character(parse_number(OppNum)))
ranks = output[c('Pair.Num', 'Rank')]
output_tidy = left_join(output_tidy, ranks, by = c('OppNum' = 'Pair.Num'))

output_tidy$OppNum = output[output$Pair.Num == output_tidy$OppNum, 'Rank']
avg_opp_rank = output_tidy %>%
  group_by(Player.Name) %>%
  summarize(AvgOppRank = round(mean(as.numeric(Rank.y), na.rm = T), 0))

### NOT PRODUCING CORRECT OPPONENT RANKING AVERAGES!!! ***

```


```{r warning=FALSE}
url = getURL('https://raw.githubusercontent.com/haobruce/CUNY/master/DATA607/Project1/tournamentinfo.txt')
chess = read.csv(text=url, stringsAsFactors = F)

# initialize empty data frame
df = data.frame('PlayerName' = character(), 'PlayerState'= character(), 'TotalPoints' = numeric(), 
                'PreRating' = integer(), 'OpponentAvgPreRating' = integer(), 'Opp1' = integer(), 
                'Opp2' = integer(), 'Opp3' = integer(), 'Opp4' = integer(), 'Opp5' = integer(), 
                'Opp6' = integer(), 'Opp7' = integer(), stringsAsFactors = F)  


# loop through chess txt file and add relevant data to data frame
for (i in 1:nrow(chess)) {
  # check if row contains player data
  if (!is.na(as.numeric(str_sub(chess[i,], 4, 5)))) {
    player = str_trim(str_sub(chess[i,], 9, 40))  # retrieve player name
    state = str_sub(chess[i+1,], 4, 5)  # retrieve player state
    points = as.numeric(str_sub(chess[i,], 42, 44))  # retrieve player points
    preRating = as.numeric(str_sub(chess[i+1,], 23, 26))  # retrieve player points
    opp1 = as.numeric(str_sub(chess[i,], 51, 52))  # retrieve opponent numbers to calculate avg later
    opp2 = as.numeric(str_sub(chess[i,], 57, 58))
    opp3 = as.numeric(str_sub(chess[i,], 63, 64))
    opp4 = as.numeric(str_sub(chess[i,], 69, 70))
    opp5 = as.numeric(str_sub(chess[i,], 75, 76))
    opp6 = as.numeric(str_sub(chess[i,], 81, 82))
    opp7 = as.numeric(str_sub(chess[i,], 87, 88))
    # use a temporary data frame to house each new row
    new_row = data.frame('PlayerName' = player, 'PlayerState' = state, 'TotalPoints' = points, 
                        'PreRating' = preRating, 'OpponentAvgPreRating' = 0.0, 'Opp1' = opp1, 
                        'Opp2' = opp2, 'Opp3' = opp3, 'Opp4' = opp4, 'Opp5' = opp5, 'Opp6' = opp6, 
                        'Opp7' = opp7, stringsAsFactors = F)  
    # append data to data frame
    df = rbind(df, new_row)
  }
}


# loop through data frame to add average opponent ratings
for (r in 1:nrow(df)) {
  oppRatings = NULL  # initialize empty vector
  for (c in 1:7) {
    colName = paste('Opp', c, sep = '')  # create dynamic reference to column name
    oppRow = df[r, colName]
    oppRatings = c(oppRatings, df[oppRow, 'PreRating'])  # append opponent rating to vector
  }
  df[r, 'OpponentAvgPreRating'] = round(mean(oppRatings, na.rm = T), 0)  # calculate avg rating and ignore NAs
}


# write df to csv file
#write.csv(df[, c('PlayerName', 'PlayerState', 'TotalPoints', 'PreRating', 'OpponentAvgPreRating')],
#          '/Users/brucehao/Google Drive/CUNY/git/DATA607/Project1/chess.csv')


# show top 10 rows of data frame
#knitr::kable(head(df[, c('PlayerName', 'PlayerState', 'TotalPoints', 'PreRating', 'OpponentAvgPreRating')], 10))
datatable(df[, c('PlayerName', 'PlayerState', 'TotalPoints', 'PreRating', 'OpponentAvgPreRating')])
